Genetic algorithms snake game

### 2024-05-28 article/article.tex
added the new version
added mutation v2

The primary objective of this study is to demonstrate how genetic algorithms (GAs) can be utilized to develop an agent that plays the Snake game. The Snake game, a classic arcade game, involves navigating a snake to collect food items while avoiding collisions with walls and the snake's own body. The GA is used to optimize the snake's behavior over successive generations, improving its ability to survive and collect food.
\subsection{Research question\label{sec:Research Question}}
How can we develop a genetic algorithm capable of outperforming the average score achieved by our group in the game of Snake?

@@

\section{Analysis\label{sec:Analysis}}
\subsection{Selection\label{sec:Selection}}
\subsection{Crossover\label{sec:Crossover}}
\subsection{Mutation\label{sec:Mutation}}

>>>>

The primary objective of this study is to demonstrate how genetic algorithms (GAs) can be utilized to develop an agent that plays the Snake game. The Snake game, a classic arcade game, involves navigating a snake to collect food items while avoiding collisions with walls and the snake's own body. The GA is used to optimize the snake's behavior over successive generations, improving its ability to survive and collect food.
\subsection{Research question\label{sec:Research Question}}
How can we develop a genetic algorithm that not only matches but exceeds our group's average score in the game of Snake by leveraging evolutionary strategies, including effective initialization, precise fitness evaluation, robust selection, and strategic use of crossover and mutation?

@@

\subsection{Mutation V.2 Jeff\label{sec:Mutation}}

Mutation is a critical component of genetic algorithms (GAs), serving to introduce genetic diversity into the population and prevent premature convergence to suboptimal solutions. By altering the genetic material of offspring, mutation helps maintain a healthy level of variability, ensuring the exploration of a broader solution space.

David E. Goldberg, in his foundational text \textit{Genetic Algorithms in Search, Optimization, and Machine Learning}, emphasizes the importance of mutation in maintaining genetic diversity and aiding the convergence of the algorithm to optimal solutions. He notes that mutation plays a crucial role in restoring lost or unexplored genetic material, thus enhancing the robustness of GAs \cite{goldberg1989genetic}.

Robert E. Marks, in \textit{Playing Games with Genetic Algorithms}, highlights the application of GAs in game strategies, noting that mutation is essential for evolving strategies that can adapt to dynamic environments and unforeseen challenges \cite{marks2001playing}.

\textbf{Types of Mutation:}
\begin{itemize}
    \item \textbf{Bit Flip Mutation:} Common in binary-encoded GAs, this method involves flipping one or more bits in a chromosome. Simple yet effective for maintaining genetic diversity. Suitable for early stages of evolution where broad exploration is needed.
    \item \textbf{Swap Mutation:} Involves swapping two genes within a chromosome. Particularly useful for problems like the Traveling Salesman Problem (TSP) and similar permutation-based problems.
    \item \textbf{Scramble Mutation:} Selects a subset of genes and randomly scrambles them. Maintains the overall genetic material but alters the order, fostering diversity.
    \item \textbf{Gaussian Mutation:} Applied to real-valued GAs, where a small Gaussian noise is added to the genes. Effective for fine-tuning solutions in continuous optimization problems.
\end{itemize}

\begin{quote}
\emph{"The role of mutation in GAs has been that of restoring lost or unexplored genetic material into the population to prevent the premature convergence of the GA to suboptimal solutions."} \cite{srinvas1994adaptive}
\end{quote}

\textbf{Implementing Mutation in the Snake Game GA:} To enhance the performance of our Snake game agent, we propose to experiment with various mutation techniques, monitoring their impact on the agent's ability to adapt and improve over successive generations. By fine-tuning the mutation rate and method, we aim to balance exploration and exploitation, leading to a more robust and high-performing Snake game strategy.

\section{Analysis\label{sec:Analysis}}
\subsection{Selection\label{sec:Selection}}
\subsection{Fitness\label{sec:Fitness}}
\subsection{Crossover\label{sec:Crossover}}
\subsection{Mutation\label{sec:Mutation}}
